---
title: "A GAMLSS re-analysis of an RT data set (with a focus on data's shape)"
author: "Fernando Marmolejo-Ramos"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

# Introduction

The following features GAMLSS techniques to model a data set. The data set will be from a study by Moret-Tatay ^[Moret-Tatay, C. et al. (2022). The effect of corrective feedback in basic cognitive tasks: a study in early childhood. _Children, 9_, 145. https://doi.org/10.3390/children9020145]. A short tutorial on GAMLSS can be found in Stasinopoulos et al. (2018) ^[Stasinopoulos, M. D., Rigby, R. A., & Bastiani, F. D. (2018). GAMLSS: A distributional regression approach. _Statistical Modelling, 18_(3–4), 248–273.]. 

The plan is to showcase GAMLSS techniques to model the data's shape; i.e. skewness and kurtosis. A paper that describes this approach can be found in de Bastiani et al. (in press) ^[de Bastiani, F. et al (in press). Bucket plot: a visual tool for skewness and kurtosis comparisons. _Brazilian Journal of Probability and Statistics_].

# Uploading the data

```{r, warning=FALSE, message=FALSE}
# data can be easily uploaded via the 'import dataset' tab in R studio...
library(readxl)
data <- read_excel(file.choose())


# retaining correct RTs that are between 250 and 1800 milliseconds
data2 <- subset(data, RT<1800 & RT>250 & Accuracy==1)

# ensuring all variables of interest are well
data2$Participant <- as.factor(data2$Participant)
data2$Group <- as.factor(data2$Group)
data2$Condition <- as.factor(data2$Condition)
```

# Modelling the shape of the model

First some GAMLSS models are created. The models will consider all the distribution's parameters. Specifically the Normal (N) and ExGaussian (EG) distributions are considered.

```{r, warning=FALSE, message=FALSE}
library(gamlss)
# model
mu<-as.formula(RT ~ Group * Condition + random(Participant))
sigma<-mu
nu<-mu


# two-parameters normal model

N <- gamlss(mu, 
            sigma.formula = sigma,
               family = NO,
               data=data2)

# three parameters exGaussian

EG <- gamlss(mu, 
            sigma.formula = sigma,
            nu.formula = nu,
               family = exGAUS,
               data=data2)

```
Let's examine the models's AIC values and look at the residuals.

```{r, warning=FALSE, message=FALSE}

# N and EG fits
list(normal_fit = N$aic, EG_fit = EG$aic)

# residuals

plot(N)
plot(EG)
```

The EG model gives a better fit as shown by the AIC and the residual plots. Ideally, residuals should have a mean of 0, variance = 1, skewness = 0, and kurtosis = 3.

In case a list of distributions need to be inspected, the following marginal distribution fitting can assist in doing so.

```{r, warning=FALSE, message=FALSE}

list<-fitDist(data2$RT, type='realplus')
list
```

The results indicate the four-parameters Box-Cox power exponential distribution gives the best fit. Let's check the residuals:


```{r, warning=FALSE, message=FALSE}

# four-parameters Box-Cox power exponential
tau <- mu

BCPE <- gamlss(mu, 
            sigma.formula = sigma,
            nu.formula = nu,
            tau.formula = tau,
               family = BCPE,
               data=data2)
# N and EG fits
list(BCPE_fit = BCPE$aic)

# residuals

plot(BCPE)
```

The shape of the residuals and the AIC of the BCPE were better than those of the N and EG distributions.


## Buckeplots

### Buckeplots with existing data (an example)

First off, let's illustrate these plots with existing data. In this example, only the location parameter is modelled and a P-spline is used on one of the numeric covariates.

```{r, warning=FALSE, message=FALSE}
library(gamlss.ggplots)

bccg <- gamlss(R~pb(Fl)+A+B, data=rent, family=BCCG)
no <- gamlss(R~pb(Fl)+A+B, data=rent, family=NO)

moment_bucket(bccg, no, no_bootstrap = 100)

# by a categorical covariate
moment_bucket_wrap(bccg, no, no_bootstrap = 100, xvar=rent$B) 

moment_bucket_wrap(bccg, no, no_bootstrap = 100, 
                   col_bootstrap = c('peru','red'),
                   colour_bucket = T,
                   # col_JB_test = 'pink',
                   xvar=rent$Fl)
```

Another function for bucketplots

```{r}
# function 'bp' in the gamlss package does the job
# but note the 'overlap' argument seems not to be working! (see rent$Fl)
# note this argument applies to numeric covariates

# thus a function sent by Fernanda de Bastiani is used here
source("bp.R")

bp.fernanda(list(bccg,no), xvar=~rent$Fl*rent$B, 
   show.legend = T,
   overlap = 0)
```
To fix the error in the argument `overlap` and generate more intuitive labels for the rows and columns, try the following

```{r, warning=FALSE, message=FALSE}

`Floor space (sqr mtr)` <-cut(rent$Fl, 4) 
# 4 cuts of equal length are requested 
#note this doesn't entail equal number of observations in each level

`Bathroom (0=no, 1=yes)` <- rent$B

bp.fernanda(list(bccg,no), 
   xvar=~`Floor space (sqr mtr)`*`Bathroom (0=no, 1=yes)`, 
   show.legend = T,no.bootstrap = 99,
   overlap = 0)
```


### Buckeplots for current data


```{r, warning=FALSE, message=FALSE}
# to give nicer looking labels for the rows and columns
Group <- data2$Group
Condition <- data2$Condition

# N=normal distribution, E=EG distribution, B=BCPE distribution

bp.fernanda(list(N,EG, BCPE), xvar=~Group*Condition, 
   show.legend = T, no.bootstrap = 99,
   text.to.show = c('N','E','B'),
   overlap = 0)
```


