---
title: "Reproducible research"
subtitle: "The place of probability distributions in statistical learning. A commented book review of “Distributions for modeling location, scale, and shape using GAMLSS in R” by Rigby et al. (2021)"
author: "Raydonal Ospina"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
header-includes:
- \usepackage[english]{babel}
- \usepackage[latin1]{inputenc}
- \usepackage{amsmath,amsthm}
- \usepackage{amssymb,latexsym}
- \usepackage{graphics}
- \usepackage{graphicx}
- \usepackage{amscd,eufrak}
- \usepackage{placeins}
- \usepackage{setspace}
- \usepackage{url}
- \usepackage{color}
- \setcounter{secnumdepth}{-1}
- \usepackage{pgfplots}
- \usepackage{subfigure}
- \usepackage{geometry}
- \usepackage{chngcntr}
- \onehalfspacing
- \counterwithin{figure}{section}
- \counterwithin{table}{section}
bibliography: references.bib  
csl: american-statistical-association.csl
---


```{r number_figures, echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE}
suppressPackageStartupMessages(library(knitr))
```


```{r startup, echo = FALSE, warning = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      message = FALSE,
                      fig.width  = 10,
                      fig.height = 8,
                      fig.cap = '',
                      fig.align = 'center', 
                      dpi = 72*2)#, global.par = TRUE)

options(width = 120, 
        scipen = 1e3)

# load the package
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(gridExtra)) # arrange plots of ggplot 
suppressPackageStartupMessages(library(SnowballC))
suppressPackageStartupMessages(library(ggforce))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(pals))
suppressPackageStartupMessages(library(scales))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(gamlss))
suppressPackageStartupMessages(library(gamlss.dist))
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(ggridges))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(parallel))
suppressPackageStartupMessages(library(compiler))
# get package on
# https://cran.r-project.org/src/contrib/Archive/CPHshape/CPHshape_1.0.1.tar.gz
# Install via terminal :~$ R CMD INSTALL CPHshape_1.0.1.tar.gz
suppressPackageStartupMessages(library(CPHshape))
suppressPackageStartupMessages(library(outliers))
suppressPackageStartupMessages(library(EnvStats))
suppressPackageStartupMessages(library(envoutliers))
# Install via github
# devtools::install_github('ousuga/RelDists', force=TRUE)
suppressPackageStartupMessages(library(RelDists))
# Aditional sources
source('BEorb.R')
source('BEor.R')
source('gamlssRobust.R') # Load script to robust estimation 

# My palette
my_color <- as.vector(c( polychrome(26), glasbey(32), alphabet2(26)))
#=======================================================================
# mywormplot
mywormplot <- function (m = NULL, residuals = NULL, 
                        age = NA, n.inter = 1, 
                        y.limits = c(-1, 1)) 
{
  if (inherits(m, "gamlss")) 
    residuals <- residuals(m)
  mm <- tibble::tibble(x = seq(-4, 4, l = 1000), yu = 1.96 * 
                         sqrt(stats::pnorm(.data$x) * 
                                (1 - stats::pnorm(.data$x))/length(residuals))/stats::dnorm(.data$x),
                       yl = -.data$yu)
  mm <- reshape2::melt(mm, id.var = "x")
  tmp <- data.frame(residuals = residuals, age = age)
  if (n.inter > 1) {
    if (all(is.na(age)))
      stop("intervals only possible of a vector of ages  is given")
    tmp$ag <- cut(tmp$age, n.inter)
  }
  else {
    tmp$ag <- "all ages"
  }
  tmp <- dplyr::group_by(tmp, .data$ag) %>% tidyr::nest()
  tmp <- dplyr::mutate(tmp, qq = purrr::map(.data$data, function(x) {
    qq <- as.data.frame(stats::qqnorm(x$residuals, plot.it = F))
    qq$y <- qq$y - qq$x
    qq
  }))
  tmp <- tidyr::unnest(tmp, .data$qq)
  ggplot2::ggplot(tmp, ggplot2::aes_string(x = "x", y = "y")) + 
    ggplot2::geom_point(shape = 21, size = 3, colour = "#2171B5") + 
    ggplot2::geom_line(data = mm, inherit.aes = F, 
                       ggplot2::aes_string(x = "x", 
                                           y = "value", group = "variable"), 
                       linetype = 2, size=1.2, colour = "forestgreen") + 
    ggplot2::geom_vline(xintercept = 0, colour = "firebrick3", 
                        linetype = 4, size=1.2) + ggplot2::scale_y_continuous(limits = 
                                                                                y.limits) + 
    # ggplot2::facet_wrap(~ag) + 
    ggplot2::labs(x = "Unit normal quantiles", 
                  y = "Deviation") 
     #+ ggplot2::theme_minimal() + ggplot2::theme(legend.position = "none", )
  #                                                                                                         axis.title = ggplot2::element_text(size = 13, colour = "black"), 
  #                                                                                                         axis.text = ggplot2::element_text(size = 13, colour = "black"), 
  #                                                                                                         title = ggplot2::element_text(colour = "black"))
}
```
### Robust estimation with GAMLSS: Dealing with outliers
The presence of outliers may provoke big errors in the estimation procedures [@rousseeuw2011robust;@hawkins1980identification;@maronna2019robust;@aeberhard2021robust]. Since robustness is the ability to perform well when the data obey the assumed model and to not provide completely useless results when the observations do not exactly follow it. In these example studies, we consider contaminated models, where samples $\{X_1,\ldots, X_n\}$ are identically distributed random variables.



### Example 1: Bounded data in (0, 1) (Mixture model with beta distribution)

The example use an mixture distribution to fit robust model by using GAMLSS with bounded data. The reference distribution is the Beta distribution $Beta(5,5)$ with symmetric shape around the value 0.5.
The pattern of the contamination two-side is defined as follow:
$$
\alpha_1{\cal U}(0, 0.1)+\alpha_2{\cal U}(0.9, 1)+(1-\alpha_1-\alpha_2)Beta(5,5).
$$ Here, $n=500,$ $\alpha_1=0.01=1\%$ (contamination to lower-tail), $\alpha_2=0.01=1\%$ (contamination to upper-tail).  On the contamination model we expect to generate potential extreme values that can affect the estimation of the Beta model parameters.

```{r, echo=F, warning=FALSE, message=FALSE}
# Mixture of three distributions
#creating data
# Fix seed
set.seed(31416)

# sample size
number_of_obs <- 500

# proportions of mix
proportions <- sample(c(0, 1, 2), number_of_obs,
                       replace = TRUE, prob =c(0.01, 0.01, 1-0.01-0.01))

# Reference distribution (Beta(5,5))
shape_1 <- 5
shape_2 <- 5

beta_ref <- rBEo(number_of_obs, shape_1, shape_2)

# lower tail
unif_1 <- runif(number_of_obs, 0, 0.1)
  
# upper tail
unif_2 <- runif(number_of_obs, 0.9, 1)

#
mixture_simulation <- data.frame(y = ifelse(proportions == 0,unif_1,
                                            ifelse(proportions == 1, unif_2, beta_ref)))
```

```{r,  echo = FALSE, fig.width = 10, fig.height=5, warning=FALSE, message=FALSE,  results='hide', fig.cap="Histogram (left panel) and Boxplot (rigth panel) of simulated data"}
# Histogram
p.1 <- ggplot(data = mixture_simulation) +
  geom_histogram(aes(x=y, y=..density..),bins = nclass.scott(mixture_simulation$y), colour="black", fill="white")

#Boxplot
p.2 <- ggplot(mixture_simulation, aes(x = "", y = y)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8,
             outlier.size=2, notch=TRUE,width = 0.4, colour="darkblue")+geom_jitter(colour="blue", alpha=0.2, width = 0.2)+xlab("")

out <- grid.arrange(p.1, p.2, ncol=2,   widths = c(1.5, 1.5))
return(out)
```
# Outliers detection
Now, we extract the values of the potential outliers based on the IQR criterion. According to the IQR, there are 3 outliers  

```{r,  echo = FALSE, fig.width = 10, fig.height=5, warning=FALSE, message=FALSE}
x <- seq(0.01,0.99,length=500)

#Data
y <- mixture_simulation$y

# Detection of outliers - IQR criterion
value <- boxplot.stats(y)$out
index <- which(y %in% c(value))
out_pot <- tibble("index"=index, "value"=value)
out_pot %>% kable(caption = "IQR criterion") %>% 
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, font_size = 12)
```
Now we use the Hampel [@hampel1974influence;@liu2004line] criterion for the detection of outliers. According to the Hampel filter, there are 14 outliers. 

```{r,  echo = FALSE, fig.width = 10, fig.height=5, warning=FALSE, message=FALSE}
# Bounds of Hampel criterion
lower_bound <- median(y) - 3 * mad(y, constant = 1)
upper_bound <- median(y) + 3 * mad(y, constant = 1)

# identification
index <- which(y < lower_bound | y > upper_bound)
value <- y[index]
out_pot <- tibble("index"=index, "value"=value)
out_pot %>% kable(caption = "Hampel criterion") %>% 
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, font_size = 12)
```
By the symmetry of the data, we now use the outlier informations in setting the value $k$ of potential outliers  for the Rosner's test [@rosner1983percentage].

```{r,  echo = FALSE, fig.width = 10, fig.height=5, warning=FALSE, message=FALSE}
out_3 <- rosnerTest(y, k=3, warn=FALSE)$all.stats
out_3 <- out_3[order(out_3[,5]),]
as_tibble(out_3) %>% kable(caption = "Rosner test -  k based on IQR criterion") %>% 
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, font_size = 12) %>% column_spec(c(5,8), bold = TRUE)

out_14<- rosnerTest(y, k=14, warn=FALSE)$all.stats
out_14 <- out_14[order(out_14[,5]),]
as_tibble(out_14) %>% kable(caption = "Rosner test -  k based on Hampel criterion") %>% 
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, font_size = 12) %>% column_spec(c(5,8), bold = TRUE)
```
The Rosner’s Tests show that any of observation are outlier based on the two criteria of detection. However, this result can be masked by normality violation of data. 

Finally, we apply the semiparametric method based on kernel smoothing and extreme value theory [@holevsovsky2018semiparametric; @vcampulova2018nonparametric; @holevsovsky2020estimation] for outlier detection. The outliers are identified as observations whose values are exceeded on a certain average by using the function `RDetect.outliers.EV()` of the `envoutliers` package in `R` .  According to the kernel approach, there are 9 outliers.

```{r,  echo = FALSE, fig.width = 9, fig.height=6, warning=FALSE, message=FALSE}
# Identification of outliers using extreme value theory
result = KRDetect.outliers.EV(y)
# plot outliers
plot(result)
# identification
index <- result$index[result$outlier==TRUE]  
value <- y[index]
out_pot <- tibble("index"=index, "value"=value)
out_pot %>% kable(caption = "Kernel criterion") %>% 
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, font_size = 12)
```
# GAMLSS fit
We are looking for any lawful curves in the data that may suggest that the fitted model is better or worse at predictions.  Worm-plots  [@buuren2001worm] can be used to identify some characteristics of the data that are not adequately captured by the fitted model via GAMLSS. We fit the data as a [rigby2019distributions] subsection 12.2.2. 

```{r,  echo = FALSE, fig.width = 13, fig.height=4, warning=FALSE, message=FALSE, ,  results='hide'}
# fitting BEo to contaminated data
# True
b1<-gamlss(y~1,family=BEo,n.cyc=100, trace=FALSE)
p.1 <- mywormplot(b1)+theme_bw()+ggtitle("Fit using Beta distribution")
r.1 <- cbind(fitted(b1)[1], fitted(b1,"sigma")[1])

# fit with bias correction 
mu1 <- 5
sigma1 <- 5
b2 <- gamlss(y~1,family=BEor,n.cyc=100, trace=FALSE)
r.2 <- cbind(fitted(b2)[1], fitted(b2,"sigma")[1])

#wp(b3, ylim.all=1)
p.2 <- mywormplot(b2)+theme_bw()+ggtitle("Fit using Beta distribution with bias correction")

# fit via robust estimation based on 
b3 <- gamlssRobust(b2,bound=2.878162, CD.bound=3.208707,trace=FALSE)
r.3 <- cbind(fitted(b3)[1], fitted(b3,"sigma")[1])
#wp(b3, ylim.all=1.5)
p.3 <- mywormplot(b3)+theme_bw()+ggtitle("Robust fit")

out <- grid.arrange(p.1, p.2, p.3, ncol=3)
return(out)
```
By visual inspection of the worm plots we observe that the robust fit produces the better fit of the data. 


```{r,  echo = FALSE, fig.width = 9, fig.height=6, warning=FALSE, message=FALSE}
# Estimates
Parameter <- c("True", "Fit Beta", "Fit beta with bias correction", "Robust fitted" )
mu <- c(mu1,   fitted(b1,"mu")[1],   fitted(b2,"mu")[1],    fitted(b3,"mu")[1])
sigma <- c(sigma1,fitted(b1,"sigma")[1],fitted(b2,"sigma")[1], fitted(b3,"sigma")[1])
as_tibble(cbind(Parameter, mu, sigma)) %>% kable(caption = "Fit results") %>% 
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, font_size = 12)
```
# Comparison of the fits to the contaminated beta data

```{r,  echo = FALSE, fig.width = 13, fig.height=4, warning=FALSE, message=FALSE}
ltype <- c(1,2,3,4)
colour <- c(gray(0),gray(.1),gray(.2),gray(.3),gray(.4) )

# Density
beta_dist <- data.frame(cbind(
  x, dBEo(x,mu=mu[1],sigma=sigma[1]), 
  dBEo(x,mu=mu[2],sigma=sigma[2]),
  dBEo(x,mu=mu[3],sigma=sigma[3])))
colnames(beta_dist) <- c("x","true","non_robust","robust")


beta_dist <- gather(beta_dist, distribution, pdf, true:robust)

p.1 <- ggplot(data = beta_dist) +
  geom_histogram(data = as.data.frame(y), aes(x=y, y=..density..), bins = nclass.scott(y), colour="black", fill="white")+
  geom_line(aes(x, pdf, color=distribution))+ labs(
    x=expression(italic(y)), 
    y=expression(italic(f(y))), 
    color="Beta PDF")+ theme_bw()+theme(
    legend.position = c(.03, 0.995),
    legend.justification = c("left", "top"),
    legend.box.just = "left",
    legend.margin = margin(0, 0, 0, 0)
  )
# p.1
# ggsave('betaPDF.png', p.1, width = 5.5, height = 3, units = 'in')

# CDF
beta_cdf <- data.frame(cbind(
  x, pBEo(x,mu=mu[1],sigma=sigma[1]), 
  pBEo(x,mu=mu[2],sigma=sigma[2]),
  pBEo(x,mu=mu[3],sigma=sigma[3])))
colnames(beta_cdf) <- c("x","true","non_robust","robust")


beta_cdf <- gather(beta_cdf, distribution, pdf, true:robust)

p.2 <- ggplot(data = beta_cdf) +
stat_ecdf(data = as.data.frame(y), aes(x=y), geom="point",colour="black", shape=1, size=1.5)+
  geom_line(aes(x, pdf, color=distribution))+ labs(
    x=expression(italic(y)), 
    y=expression(italic(F(y))), 
    color="Beta CDF")+ theme_bw()+theme(
    legend.position = c(.03, 0.995),
    legend.justification = c("left", "top"),
    legend.box.just = "left",
    legend.margin = margin(0, 0, 0, 0)
  )
# p.2
# ggsave('betaCDF.png', p.2, width = 5.5, height = 3, units = 'in')

n <- length(y)
    t <- order(y)
    sortx <- x[t]
    Trn <- rep(NA, times = n)
    r <- rep(NA, times = n)
    Trn[1] <- n * sortx[1]
    r[1] <- 1/n
    for (i in 2:n) {
        Trn[i] <- Trn[i - 1] + (n - i + 1) * (sortx[i] - sortx[i - 
            1])
        r[i] <- i/n
    }
    
    Trn <- sort(Trn)
# hazard
beta_hzd <- data.frame(cbind(
  x, 
  dBEo(x,mu=mu[1],sigma=sigma[1])/(1-pBEo(x,mu=mu[1],sigma=sigma[1])), 
  dBEo(x,mu=mu[2],sigma=sigma[2])/(1-pBEo(x,mu=mu[2],sigma=sigma[2])),
  dBEo(x,mu=mu[3],sigma=sigma[3])/(1-pBEo(x,mu=mu[3],sigma=sigma[3]))))
colnames(beta_hzd) <- c("x","true","non_robust","robust")


beta_hzd <- gather(beta_hzd, distribution, pdf, true:robust)

emp_hzd <- find.shapeMLE(y,  type = "increasing")
H <- exp(emp_hzd$H) 

p.3 <- ggplot(data = beta_hzd) +
 # geom_point(data = as.data.frame(H), aes(x=x, y=H), colour="black", shape=1, size=1.8)+
  # geom_point(data = as.data.frame(Trn), aes(x=x, y=Trn), colour="blue", shape=1, size=1.8)+
  geom_line(aes(x, pdf, color=distribution))+ labs(
    x=expression(italic(y)), 
    y=expression(italic(h(y))), 
    color="Beta Hazard")+ theme_bw()+theme(
    legend.position = c(.03, 0.995),
    legend.justification = c("left", "top"),
    legend.box.just = "left",
    legend.margin = margin(0, 0, 0, 0)
  )
# p.3
# ggsave('betaHazard.png', p.3, width = 5.5, height = 3, units = 'in')

# df <- as.data.frame(y)
# p <-
#   ggplot(df, aes(x=y)) +
#   stat_ecdf(geom="point",colour="black", shape=1, size=1.5)+
#   stat_function(fun=pBEo, colour="#00AFBB", size=1, args = list(mu=mu[1],sigma=sigma[1])) +
#   stat_function(fun=pBEo, colour= "#E7B800", size=1, args = list(mu=mu[2],sigma=sigma[2]))+ 
#   stat_function(fun=pBEo, colour="#FC4E07", size=1, args = list(mu=mu[3],sigma=sigma[3])) +
#   labs(x="", y="")+xlim(0.01,0.99)
out <- grid.arrange(p.1, p.2, p.3, ncol=3)
```

### Example 2: Reaction time distributions

The example uses a mixture distribution to generate contaminated Reaction times. We fit known different Reaction time distribution and use robust estimation via GAMLSS. The reference distribution is the ExGaussian distribution $ExGaussian(\mu,\sigma,\nu).$ The parameters $\mu$ and $\sigma$ are the mean and standard deviation from the normal distribution variable while the parameter $\nu$ is the rate control of the exponential component (the right-skew of the distribution).

The pattern of the contamination two-side is defined as follow:
$$
(1-\alpha)ExGaussian(\mu,\sigma,\nu)+\alpha{\cal U}(L,U).
$$ Here, $\alpha$ is the contamination fraction, $L$ is the smallest observed reaction time, and $U$ is the largest. 
On the contamination model we expect to generate potential extreme values that can affect the estimation of the ExGaussian parameters. In this example, $n=500,$ $\alpha=0.1=10\%$ (contamination parameter), $\mu=0.5,$ $\sigma=0.1,$ $\nu = 3,$ $L=0$ and $U=5.$ 

```{r,  echo = FALSE, fig.width = 13, fig.height=4, warning=FALSE, message=FALSE}

# Mixture of three distributions
#creating data
# Fix seed
set.seed(31416)

# sample size
number_of_obs <- 500

# proportions of mix
proportions <- sample(c(1, 0), number_of_obs,
                       replace = TRUE, prob =c(0.9, 0.1))

# Reference ExGaussian(.5,0.1,3)
mu <- 0.5
sigma <- 0.1
nu <- 3

exgauss_ref <- gamlss.dist::rexGAUS(number_of_obs, mu=mu, sigma = sigma, nu = nu)

# Tail (outliers)
unift <- runif(number_of_obs, 0, 10)
  
# response
mixture_simulation <- data.frame(y = ifelse(proportions == 1,exgauss_ref, unift))
```

```{r,  echo = FALSE, fig.width = 10, fig.height=5, warning=FALSE, message=FALSE,  results='hide', fig.cap="Histogram (left panel) and Boxplot (rigth panel) of simulated data"}
# Histogram
p.1 <- ggplot(data = mixture_simulation) +
  geom_histogram(aes(x=y, y=..density..),bins = nclass.scott(mixture_simulation$y), colour="black", fill="white")

#Boxplot
p.2 <- ggplot(mixture_simulation, aes(x = "", y = y)) +
  geom_boxplot(outlier.colour="red", outlier.shape=8,
             outlier.size=2, notch=TRUE,width = 0.4, colour="darkblue")+geom_jitter(colour="blue", alpha=0.2, width = 0.2)+xlab("")

out <- grid.arrange(p.1, p.2, ncol=2,   widths = c(1.5, 1.5))
return(out)
```
# Outlier detection
Here, according to the kernel approach, there are 9 outliers.

```{r,  echo = FALSE, fig.width = 9, fig.height=6, warning=FALSE, message=FALSE}
# Identification of outliers using extreme value theory
y <- mixture_simulation$y
result = KRDetect.outliers.EV(y)
# plot outliers
plot(result)
# identification
index <- result$index[result$outlier==TRUE]  
value <- y[index]
out_pot <- tibble("index"=index, "value"=value)
out_pot %>% kable(caption = "Kernel criterion") %>% 
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, font_size = 12)
```

We fit the Exgaussian, Lindley, Gamma and Weibull  models for data using GAMLSS. Also, ExGaussian is robust fitted by use `gamlssRobust`. The generalized joint regression modelling implemented in the `GJRM` package of `R` for robus estimation produce wrong results, we decided not to use it. Here, the Lindley distribution is in `RelDists` package. [See here](https://github.com/ousuga/RelDists).
```{r,  echo = FALSE, fig.width = 13, fig.height=9, warning=FALSE, message=FALSE,   results='hide'}
# fitting Exgaussian to contaminated data
# ExGaussian
b1<-gamlss(y~1,family=exGAUS,n.cyc=100, trace=FALSE)
p.1 <- mywormplot(b1)+theme_bw()+ggtitle("ExGaussian")
r.1 <- cbind(fitted(b1)[1], fitted(b1,"sigma")[1],fitted(b1,"nu")[1])

# ExGaussian fit via robust estimation based  
b2 <- gamlssRobust(b1,bound=2.878162, CD.bound=3.208707,trace=FALSE)
r.2 <- cbind(fitted(b2)[1], fitted(b2,"sigma")[1],fitted(b2,"nu")[1]) 
p.2 <- mywormplot(b2)+theme_bw()+ggtitle("Robust ExGaussian")


# Lindley - Based on RelDists package
b3<-gamlss(y~1,family=LIN,n.cyc=100, trace=FALSE)
p.3 <- mywormplot(b3)+theme_bw()+ggtitle("Lindley")
r.3 <- cbind(fitted(b3)[1])

# Gamma
b4<-gamlss(y~1,family=GA,n.cyc=100, trace=FALSE)
p.4 <- mywormplot(b4)+theme_bw()+ggtitle("Gamma")
r.4 <- cbind(fitted(b4)[1], fitted(b4,"sigma")[1])
 

# IG - Inverse Gaussian
b5<-gamlss(y~1,family=IG,n.cyc=100, trace=FALSE)
p.5 <- mywormplot(b5)+theme_bw()+ggtitle("Inverse Gaussian")
r.5 <- cbind(fitted(b5)[1], fitted(b5,"sigma")[1])

# Weibull
b6<-gamlss(y~1,family=IG,n.cyc=100, trace=FALSE)
p.6 <- mywormplot(b6)+theme_bw()+ggtitle("Weibull")
r.6 <- cbind(fitted(b6)[1], fitted(b6,"sigma")[1])


# Fit robust estimation via GJRM - wrong estimation processes
# I am trying different approaches, but the results are poor or simply the algorithm not converge
# eq.mu <- y~1
# eq.s <- ~1
# fl <- list(eq.mu,eq.s)
# br<- GJRM::gamlss(fl, margin ="WEI", robust=TRUE, data=mixture_simulation)
# r.r <- cbind(fitted(b5r)[1], fitted(br,"sigma")[1])
# p.r <- mywormplot(br)+theme_bw()+ggtitle("Robust ExGaussian via GJRM")

out.1 <- grid.arrange(p.1, p.2, p.3, p.4,p.5,p.6,  ncol=3)
```
By visual inspection of the worm plots we observe that the robust fit produces the better fit of the data.


```{r,  echo = FALSE, fig.width = 9, fig.height=6, warning=FALSE, message=FALSE,  results='hide'}
# Estimates
Parameter <- c("True", "ExGauss", "Robust ExGauss", "Lindley", "Gamma", "IG", "Weibull" )

mu.1 <- c(mu,   fitted(b1,"mu")[1],   fitted(b2,"mu")[1],    fitted(b3,"mu")[1], fitted(b4,"mu")[1],   fitted(b5,"mu")[1],    fitted(b6,"mu")[1] )

sigma.1 <- c(sigma,   fitted(b1,"sigma")[1],   fitted(b2,"sigma")[1],  "",   fitted(b4,"sigma")[1],    fitted(b5,"sigma")[1], fitted(b6,"sigma")[1]  )

nu.1 <- c(nu,   fitted(b1,"nu")[1],   fitted(b2,"nu")[1], "",   "",    "", "")

as_tibble(cbind(Parameter, mu.1, sigma.1, nu.1)) %>% kable(caption = "Fit results") %>% 
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE, font_size = 12)
```


```{r,  echo = FALSE, fig.width = 13, fig.height=4, warning=FALSE, message=FALSE,  results='hide'}
ltype <- c(1,2,3,4)
colour <- c(gray(0),gray(.1),gray(.2),gray(.3),gray(.4) )

 x=seq(min(y),max(y), length=500)

# Density
ExGa_dist <- data.frame(cbind(
  x, 
  dexGAUS(x,mu=fitted(b1,"mu")[1],sigma=fitted(b1,"sigma")[1],nu=fitted(b1,"nu")[1]), 
  dexGAUS(x,mu=fitted(b2,"mu")[1],sigma=fitted(b2,"sigma")[1],nu=fitted(b2,"nu")[1]),
  dLIN(x,mu=fitted(b3,"mu")[1]),
  dGA(x, mu=fitted(b4,"mu")[1],sigma=fitted(b4,"sigma")[1]),
  dIG(x, mu=fitted(b5,"mu")[1],sigma=fitted(b5,"sigma")[1]),
  dWEI(x,mu=fitted(b6,"mu")[1],sigma=fitted(b6,"sigma")[1])
  ))


colnames(ExGa_dist) <- c("x","ExGauss","Robust ExGauss", "LIN", "GA", "IG", "WEI")


ExGa_dist <- gather(ExGa_dist, distribution, pdf, ExGauss:WEI)

p.1 <- ggplot(data = ExGa_dist) +
  geom_histogram(data = as.data.frame(y), aes(x=y, y=..density..), bins = nclass.scott(y), colour="black", fill="white")+
  geom_line(aes(x, pdf, color=distribution))+ labs(
    x=expression(italic(y)), 
    y=expression(italic(f(y))), 
    color="PDF")+ theme_bw()+theme(
    legend.position = c(.6, 0.995),
    legend.justification = c("left", "top"),
    legend.box.just = "left",
    legend.margin = margin(0, 0, 0, 0)
  )

# ggsave('betaPDF.png', p.1, width = 5.5, height = 3, units = 'in')

#CDF
ExGa_cdf <- data.frame(cbind(x,
  pexGAUS(x,mu=fitted(b1,"mu")[1],sigma=fitted(b1,"sigma")[1],nu=fitted(b1,"nu")[1]), 
  pexGAUS(x,mu=fitted(b2,"mu")[1],sigma=fitted(b2,"sigma")[1],nu=fitted(b2,"nu")[1]),
  pLIN(x,mu=fitted(b3,"mu")[1]),
  pGA(x, mu=fitted(b4,"mu")[1],sigma=fitted(b4,"sigma")[1]),
  pIG(x, mu=fitted(b5,"mu")[1],sigma=fitted(b5,"sigma")[1]),
  pWEI(x,mu=fitted(b6,"mu")[1],sigma=fitted(b6,"sigma")[1])
  ))
colnames(ExGa_cdf) <- c("x","ExGauss","Robust ExGauss", "LIN", "GA", "IG", "WEI")



ExGa_cdf <- gather(ExGa_cdf, distribution, cdf, ExGauss:WEI)

p.2 <- ggplot(data = ExGa_cdf) +
stat_ecdf(data = as.data.frame(y), aes(x=y), geom="point",colour="black", shape=1, size=1.5)+
  geom_line(aes(x, cdf, color=distribution))+ labs(
    x=expression(italic(y)),
    y=expression(italic(F(y))),
    color="Beta CDF")+ theme_bw()+theme(
    legend.position = c(.6, 0.5),
    legend.justification = c("left", "top"),
    legend.box.just = "left",
    legend.margin = margin(0, 0, 0, 0)
  )


# hazard
ExGa_hzd <- data.frame(cbind(
  x,
  dexGAUS(x,mu=fitted(b1,"mu")[1],sigma=fitted(b1,"sigma")[1],nu=fitted(b1,"nu")[1])/(1-pexGAUS(x,mu=fitted(b1,"mu")[1],sigma=fitted(b1,"sigma")[1],nu=fitted(b1,"nu")[1])), 
  
  dexGAUS(x,mu=fitted(b2,"mu")[1],sigma=fitted(b2,"sigma")[1],nu=fitted(b2,"nu")[1])/(1-pexGAUS(x,mu=fitted(b2,"mu")[1],sigma=fitted(b2,"sigma")[1],nu=fitted(b2,"nu")[1])),
  
  dLIN(x,mu=fitted(b3,"mu")[1])/(1-pLIN(x,mu=fitted(b3,"mu")[1])),
  
  dGA(x, mu=fitted(b4,"mu")[1],sigma=fitted(b4,"sigma")[1])/(1-pGA(x, mu=fitted(b4,"mu")[1],sigma=fitted(b4,"sigma")[1])),
  
  dIG(x, mu=fitted(b5,"mu")[1],sigma=fitted(b5,"sigma")[1])/(1-pIG(x, mu=fitted(b5,"mu")[1],sigma=fitted(b5,"sigma")[1])),
  
  dWEI(x,mu=fitted(b6,"mu")[1],sigma=fitted(b6,"sigma")[1])/(1-pWEI(x,mu=fitted(b6,"mu")[1],sigma=fitted(b6,"sigma")[1]))
  
  ))


colnames(ExGa_hzd) <-  c("x","ExGauss","Robust ExGauss", "LIN", "GA", "IG", "WEI")


ExGa_hzd <- gather(ExGa_hzd, distribution, hzd, ExGauss:WEI)

emp_hzd <- find.shapeMLE(y,  type = "increasing")
H <- exp(emp_hzd$H)

p.3 <- ggplot(data = ExGa_hzd) +
 # geom_point(data = as.data.frame(H), aes(x=x, y=H), colour="black", shape=1, size=1.8)+
  # geom_point(data = as.data.frame(Trn), aes(x=x, y=Trn), colour="blue", shape=1, size=1.8)+
  geom_line(aes(x, hzd, color=distribution))+ labs(
    x=expression(italic(y)),
    y=expression(italic(hzd(y))),
    color="Hazard")+ theme_bw()+theme(
    legend.position = c(.6, 0.995),
    legend.justification = c("left", "top"),
    legend.box.just = "left",
    legend.margin = margin(0, 0, 0, 0)
  )

# ggsave('betaHazard.png', p.3, width = 5.5, height = 3, units = 'in')

# df <- as.data.frame(y)
# p <-
#   ggplot(df, aes(x=y)) +
#   stat_ecdf(geom="point",colour="black", shape=1, size=1.5)+
#   stat_function(fun=pBEo, colour="#00AFBB", size=1, args = list(mu=mu[1],sigma=sigma[1])) +
#   stat_function(fun=pBEo, colour= "#E7B800", size=1, args = list(mu=mu[2],sigma=sigma[2]))+ 
#   stat_function(fun=pBEo, colour="#FC4E07", size=1, args = list(mu=mu[3],sigma=sigma[3])) +
#   labs(x="", y="")+xlim(0.01,0.99)
out <- grid.arrange(p.1, p.2, p.3, ncol=3)
out
```

# References

<div id="refs"></div>

